---
title: "Calibration Dashboard - San Ysidro AB617"
runtime: shiny
author:
  - name: "Elena Austin"
    # email: elaustin@uw.edu
    # affiliation: University of Washington
    # footnote: Corresponding Author
  - name: "Edmund Seto"
    # email: eseto@uw.edu
    # affiliation: University of Washington
output: 
  flexdashboard::flex_dashboard:
    navbar:
      - { icon: "fa-question-circle", href: "https://deohs.washington.edu/san-ysidro-air-quality-and-border-traffic-study", align: right }
    orientation: columns
    source_code: "https://github.com/elaustin/calibration_dashboard"
    vertical_layout: fill
    theme: readable
    
---

```{r setup, include=FALSE}

library(ggplot2)
library(data.table)
library(shiny)
library(ggthemes)
library(DT)
library(flexdashboard)
library(stargazer)
library(lubridate)
library(lme4)

options(shiny.maxRequestSize = 30*1024^2)
```

Inputs {.sidebar data-width=300}
===============================

Select Your Data Files

```{r}
fileInput("file_sensor", "Choose Sensor data file (CSV)",
                multiple = TRUE,
                accept = c("text/csv",
                         "text/comma-separated-values,text/plain",
                         ".csv"))

fileInput("file_ref", "Choose Reference data file (CSV)",
                multiple = TRUE,
                accept = c("text/csv",
                         "text/comma-separated-values,text/plain",
                         ".csv"))
fileInput("file_pred", 
          "Choose prediction data file (csv)",
                multiple = TRUE,
                accept = c("text/csv",
                         "text/comma-separated-values,text/plain",
                         ".csv"))

 actionButton("submit", "Submit")
 
renderUI({selectizeInput("time_var", label = "Time",
        choices = colnames(reference()), selected = "datetime")})

renderUI({selectizeInput("ref_var", label = "REFERENCE",
        choices = colnames(reference()), selected = "PM25HR")})

renderUI({selectizeInput("sens_var", label = "SENSOR",
        choices = colnames(sensor()), selected = "pm25mean")})

renderUI({selectizeInput("rh_var", label = "Control Variable (e.g. RH)",
        choices = c("No RH",
                    colnames(sensor())), selected = "No RH")})

renderUI({selectizeInput("temp_var", label = "Second Control Variable (e.g. T)",
        choices = c("No Temperature",
                    colnames(sensor())), selected = "No Temperature")})

renderUI({selectizeInput("id_var", label = "Monitor ID",
        choices = c("No Monitor ID",
                    colnames(sensor())), selected = "No Monitor ID")})

actionButton("calibrate", "Calibrate")
 

```


Regression
===================================== 

Column {.tabset}
-----------------------------------------------------------------------

### Calibration Summary

```{r}

reference <- eventReactive(input$submit, {
  req(input$file_ref)
  fread(input$file_ref[["datapath"]])
})

sensor <- eventReactive(input$submit, {
  req(input$file_sensor)
  fread(input$file_sensor[["datapath"]])
})

pred <- eventReactive(input$submit, {
  req(input$file_pred)
  fread(input$file_pred[["datapath"]])
})

#This section can be used to customize the dashboard code  
calib <- eventReactive(input$calibrate 
                       & input$submit, {
  
  all = dataset()
  
  #no explanatory variable
  if(input$rh_var %in% "No RH")
  {
    return(lm( all[[input$ref_var]]  ~ 
               all[[input$sens_var]], 
       na.action = "na.exclude"))
  }
  
  #adjusting for 2 explanatory variable
  if(!(input$rh_var %in% c("No RH")) & (input$temp_var %in% c("No Temperature")))
  {
  return(lm( all[[input$ref_var]]  ~ 
               all[[input$sens_var]] + 
       all[[input$rh_var]] + 
         I(all[[input$rh_var]]^2), 
       na.action = "na.exclude"))
  }
  
  #adjusting for 2 explanatory variable
  if(!(input$rh_var %in% c("No RH")) & !(input$temp_var %in% c("No Temperature")))
  {
  return(lm( all[[input$ref_var]]  ~ 
               all[[input$sens_var]] + 
       all[[input$rh_var]] + I(all[[input$rh_var]]^2) + all[[input$temp_var]] + I(all[[input$temp_var]]^2), 
       na.action = "na.exclude"))
  }
  
  #regression by monitor
  if(!(input$id_var %in% "No Monitor ID"))
  {
  all$randomv = all[[input$id_var]]
  return(lmer( all[[input$ref_var]]  ~ 
               all[[input$sens_var]] + 
       all[[input$rh_var]] + 
         I(all[[input$rh_var]]^2) + 
         all[[input$temp_var]] + 
         I(all[[input$temp_var]]^2) +
       (1|randomv),
       data = all,
       na.action = "na.exclude"))
    }
  
})

dataset <- eventReactive(input$calibrate, {
  refdata = reference()
  sendata = sensor()
  refdata[, datetime := parse_date_time(get(input$time_var), 
                                        orders = c("YmdHMS", "YmdHM","mdYHMS","mdYHM"))]
  sendata[, datetime := parse_date_time(get(input$time_var), 
                                        orders = c("YmdHMS", "YmdHM","mdYHMS", "mdYHM"))]
  #average 1-hour
  sendata[, hourly := as.POSIXct(floor_date(datetime, "1 hour"))]
  
  if(input$id_var %in% "No Monitor ID")
  {
  sendata = sendata[,
                   lapply(.SD, mean, na.rm=T), 
                   #Every hour by monitor and site name
                   by = c("hourly"),
                   #identify numeric columns
                   .SDcols = colnames(sendata)[unlist(sendata[, 
                                lapply(.SD, is.numeric)])] ]
  } else {
  sendata = sendata[,
                   lapply(.SD, mean, na.rm=T), 
                   #Every hour by monitor and site name
                   by = c("hourly", input$id_var),
                   #identify numeric columns
                   .SDcols = colnames(sendata)[unlist(sendata[, 
                                lapply(.SD, is.numeric)])] ]
  }


  setkey(refdata, datetime)
  setkey(sendata, hourly)
  
  return(sendata[refdata])
})

dataset_pred <- eventReactive(input$calibrate, {
  
  if(!is.null(pred())){
  sendata = pred()
  sendata$calibrated.data = predict(calib(), newdata = sendata, na.action="na.exclude")
  #sendata[, hourly := as.POSIXct(hourly)]
  sendata = sendata[!is.na(calibrated.data)]
  
  return(sendata)
  }
  

  if(is.null(sendata))
    return(plotdata())
}
)





plotData <- eventReactive(input$calibrate, {
  plotdata = dataset()
  plotdata$calibrated.data = predict(calib())
  plotdata[, hourly := as.POSIXct(hourly)]
  plotdata = plotdata[!is.na(calibrated.data)]
  plotdata
})

dtData <- eventReactive(input$calibrate, {
  if(!exists(input$file_pred))
  {
  plotdata = dataset()
  plotdata$calibrated.data = predict(calib())
  plotdata[, hourly := as.POSIXct(hourly)]
  plotdata = plotdata[!is.na(calibrated.data)]
  plotdata
  }
   if(exists(input$file_pred))
   {
  plotdata = dataset_pred()
  plotdata$calibrated.data = predict(calib())
  plotdata[, hourly := as.POSIXct(hourly)]
  plotdata = plotdata[!is.na(calibrated.data)]
  plotdata
   }
})


covariate.values <- eventReactive(input$calibrate, {
 if(input$rh_var %in% "No RH"){
                          return(c("SENSOR"))
 }  else if(input$temp_var %in% "No Temperature")  
   {
     vals = c("SENSOR", "RH", "RH Squared")
     return(vals)
   } else
  {
     vals = c("SENSOR", "RH", "RH Squared", "T", "T squared")
     return(vals)
   } 
     
     
}
)

renderUI(

  HTML(
    stargazer(calib(),
              omit.stat = "adj.rsq",
                 covariate.labels = covariate.values(),
                        type = "html")))



```

Row 
-------------------------------------

### Before Calibration

```{r}

renderPlot({
  plotdata = plotData()
  
  plot(plotdata[[input$ref_var]], plotdata[[input$sens_var]],
       xlab = "Regulatory Data", ylab = "Raw Sensor Data")
  abline(a = 0, b = 1, lty =1, col = "red")

}
  )

```

### After Calibration

```{r}

renderPlot({
  
  plotdata = plotData()

  plot(plotdata[[input$ref_var]], plotdata$calibrated.data,
       xlab = "Regulatory Data", ylab = "Calibrated Sensor Data")
  abline(a = 0, b = 1, lty =1, col = "red")

}
  )

```

Calibrated Data Time Series
===================================== 

```{r}

renderPlot({
  
  plotdata = plotData()
  
  ggplot(plotdata, aes(hourly, calibrated.data, color = "Calibrated Data")) + geom_point() + 
    geom_point(data = plotdata, aes(hourly,get(input$ref_var), 
                                    color = "Reference Data")) + 
    geom_line() + 
    xlab("Date") + 
    ylab("Pollutant") +
    theme_light(16) +
    theme(legend.title = element_blank())
    

}
  )

```

Get Calibrated Data
===================================== 

```{r}


DT::renderDataTable(server = FALSE, {
    datatable( data = dataset_pred()
                , extensions = 'Buttons'
                , options = list( 
                  dom = "Blfrtip"
                  , buttons = 
                    list("copy", list(
                      extend = "collection"
                      , buttons = c("csv", "excel", "pdf")
                      , text = "Download"
                    ) ) # end of buttons customization
                   
                   # customize the length menu
                  , lengthMenu = list( c(10, 20, -1) # declare values
                                       , c(10, 20, "All") # declare titles
                  ) # end of lengthMenu customization
                  , pageLength = 10
                   
                   
                ) # end of options
               
     )} # end of datatables
   )
```

Help
===================================== 
<h4><font color="blue">About this Dashboard</font><h4>

Welcome to the low-cost sensor calibration dashboard. This page will assist you in calibrating your low-cost sensor data to a reference or regulatory instrument. 

<h4><font  color="blue">Before you begin</font><h4>

To use this page you will need to have 2 datasets ready to access on your computer.

1. The first dataset _Sensor Data_ will include the following columns:

- **timestamp**: A column containing the date and time (see _Note about dates_ for more details)
- **sensor**: A column containing the sensor measurements
- **RH**: A column with "Relative Humidity" (RH) measures (optional)
- **ID**: A column with individual sensor ID values (optional). This allows for the calibration of multiple sensors simultaneously. 


2. The second dataset <em>Reference Data<em> will include the following columns:

- **timestamp**: A column containing the date and time (see _Note about dates_ for more details)
- **reference**: A column with the reference instrument measurement.

<h4><font color="blue">Calibrating deployment measurements (optional)</font><h4>

If you retain your _Sensor Data_ and _Reference Data_ files, you can use this page to calibrate your collected sensor data after deployment. To do so, enter _Prediction Data_ file. The column names in this dataset must exactly match those used in the _Sensor Data_ file. The results can be downloaded using the **Get Calibrated Data** tab.  

The **timestamp** column in both the "Sensor Data" file and the "Reference Data" file must have exactly the same name (case-sensitive). Both datasets must be reported in the same timezone. The format of the timestamp can be any of the following: 

- YYYY-mm-dd HH:MM:SS (ex. 2019-01-15 22:00:00)
- mm/dd/YYYY HH:MM (ex. 01/15/2019 22:00)
- mm/dd/YYYY HH:MM:SS (ex. 01/15/2019 22:00:00)



